---
title: "Impact of fishing pressure on California Spiny Lobster abundance and community size structure along the Santa Barbara Coast"
author: "Shuhan Song"
date: "November 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#loading packages
library(tidyverse)
library(janitor)
library(tidyr)
library(DT)
library(plotly)

# read in csv file
lobster_abundance <- read_csv("lobster_abundance_sbc_lter.csv", na = "-99999") %>% 
  clean_names() %>% 
  uncount(count)
```

### Introduction

### Data and Methods

annual abundance summed up monthly counts of lobster at each site each year, from 2012 to 2018. 

### Results  
#### A. Annual abundance of lobsters at each site
The changes in annual abundance of California Spiny Lobster at five sites were plotted to visualize (Figure 1). Comparing 2018 to 2012, spiny lobsters from all sites experienced an increasing trend in abundance except those from Arroyo Quemado, whose stock mainly stayed the same. Populations from Marine Protected Areas (MPA) both peaked in 2015 then quickly recovered from 2016 to 2018, especially for Isla Vista lobsters. However, those from non-MPAs all declined after the peak and failed to recover to the peak level in the subsequent years.

```{r}
mpa_site <- c("NAPL", "IVEE")

abdn_year_site <- lobster_abundance %>% 
  count(year, site) %>% 
  mutate(mpa = ifelse(site %in% mpa_site, "(MPA)", "(non-MPA)")) %>% 
  mutate(year_abb = year - 2000) %>% 
  mutate(site_labels = factor(site, levels = c("AQUE", "CARP", "MOHK", "IVEE", "NAPL"),
                           labels = c("Arroyo Quemado", "Carpinteria",  "Mohawk", "Isla Vista", "Naples")))

ggplot(data = abdn_year_site, 
       aes(x = year_abb,
           y = n)) +
  geom_line(aes(color = mpa), 
            show.legend = FALSE) +
  geom_point(aes(color = mpa),
             alpha = 0.5,
             show.legend = FALSE) +
  facet_wrap(site_labels ~ mpa) +
  labs(x = "Year", 
       y = "Annual abundance") +
  theme_bw()

```

_**Figure 1.** The changes in annual abundance of spiny lobsters at each site (2012-2018). Sites were from both MPAs (red) and non-MPAs (blue). Data: SBC LTER_ 

--------------------
#### B. 
```{r}

```


#### C. 

### Summary

### References

