---
title: "Impact of fishing pressure on California Spiny Lobster abundance and community size structure along the Santa Barbara Coast"
author: "Shuhan Song"
date: "November 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#loading packages
library(tidyverse)
library(janitor)
library(tidyr)
library(DT)
library(plotly)

# read in csv file and clean the data frame
lobster_abundance <- read_csv("lobster_abundance_sbc_lter.csv", na = "-99999") %>% 
  clean_names() %>% 
  uncount(count)
```

### Introduction

### Data and Methods

annual abundance summed up monthly counts of lobster at each site each year, from 2012 to 2018. 

### Results  
#### A. Annual abundance of lobsters at each site
The changes in annual abundance of California Spiny Lobster at five sites were plotted to visualize (Figure 1). Comparing 2018 to 2012, spiny lobsters from all sites experienced an increasing trend in abundance except those from Arroyo Quemado, whose stock mainly stayed the same. Populations from Marine Protected Areas (MPA) both peaked in 2015 then quickly recovered from 2016 to 2018, especially for Isla Vista lobsters. However, those from non-MPAs all declined after the peak and failed to recover to the peak level in the subsequent years.

```{r}
mpa_site <- c("NAPL", "IVEE")

abdn_year_site <- lobster_abundance %>% 
  count(year, site) %>% 
  mutate(mpa = ifelse(site %in% mpa_site, "(MPA)", "(non-MPA)")) %>% 
  mutate(year_abb = year - 2000) %>% 
  mutate(site_labels = factor(site, levels = c("AQUE", "CARP", "MOHK", "IVEE", "NAPL"),
                           labels = c("Arroyo Quemado", "Carpinteria",  "Mohawk", "Isla Vista", "Naples")))

ggplot(data = abdn_year_site, 
       aes(x = year_abb,
           y = n)) +
  geom_line(aes(color = mpa), 
            show.legend = FALSE) +
  geom_point(aes(color = mpa),
             alpha = 0.5,
             show.legend = FALSE) +
  facet_wrap(site_labels ~ mpa) +
  labs(x = "Year", 
       y = "Annual abundance") +
  theme_light()

```

_**Figure 1.** The changes in annual abundance of spiny lobsters at each site (2012-2018). Sites were from both MPAs (red) and non-MPAs (blue). Data: SBC LTER_ 

--------------------

#### B. The shift of lobster size distribution at each site
Lobster size distribution presented different levels of shifts at each site comparing 2018 to 2012 (Figure 2). Overall, based on the observations, the size structure of lobster population shifted to the right from 2012 to 2018 with 2018 also had much greater population. Specifically, the shift to larger sizes was more significant in sites located in MPA. Lobsters in Isla Vista and Naples were clustered to smaller sizes in 2012 but grew larger in 2018 under protections from fishing pressure. However, for those in non-MPAs, lobsters did not grow larger if not dropped to smaller sizes.

```{r}
size_1218_site <- lobster_abundance %>% 
  filter(year %in% c("2012", "2018")) %>% 
  select(year, site, size_mm) %>% 
  mutate(mpa = ifelse(site %in% mpa_site, "(MPA)", "(non-MPA)")) %>% 
  mutate(site_labels = factor(site, levels = c("AQUE", "CARP", "MOHK", "IVEE", "NAPL"),
                           labels = c("Arroyo Quemado", "Carpinteria",  "Mohawk", "Isla Vista", "Naples")))

size_1218 <- size_1218_site %>% 
  transform(site_labels = NULL, mpa = NULL)

ggplot(data = size_1218_site, 
       aes(x = size_mm)) +
  geom_histogram(data = size_1218, 
                 fill = "gray90", 
                 binwidth = 5) +
  geom_histogram(aes(fill = mpa), 
                 binwidth = 5,
            show.legend = FALSE) +
  facet_grid(year ~ site_labels + mpa) +
  labs(x = "Size(mm)", 
       y = "Count") +
  theme_light()
```
_**Figure 2.** The shift of loster size distribution in 2012 and 2018 by site. In 2012, sites in non-MPA (blue) had more larger sizes lobsters than sites in MPA (red). In 2018, lobsters in MPAs had larger sizes than those from non-MPAs. Data: SBC LTER_ 

-----------------------

#### C. 
```{r}
size_1218_mpa <- lobster_abundance %>% 
  filter(year %in% c("2012", "2018")) %>% 
  select(year, site, size_mm) %>% 
  mutate(mpa = ifelse(site %in% mpa_site, "MPA", "non-MPA"))

size_table <- size_1218_mpa %>% 
  group_by(year, mpa) %>% 
  summarize(mean = round(mean(size_mm), 2),
            standard_deviation = round(sd(size_mm), 2),
            sample_size = length(size_mm))
```

_**Table 1.** Summary of spiny lobster size for MPA and non-MPA sites in 2012 and 2018. Data: SBC LTER_ 
```{r}
datatable(size_table, 
          colnames = c("Year", "MPA", "Mean", "Standard Deviation", "Sample Size"))
```

### Summary

### References

