---
title: "Impact of fishing pressure on California Spiny Lobster abundance and community size structure along the Santa Barbara Coast"
author: "Shuhan Song"
date: "November 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#loading packages
library(tidyverse)
library(janitor)
library(tidyr)
library(kableExtra)
library(effsize)

# read in csv file and clean the data frame
lobster_abundance <- read_csv("lobster_abundance_sbc_lter.csv", na = "-99999") %>% 
  clean_names() %>% 
  uncount(count)
```

### Introduction

### Data and Methods

annual abundance summed up monthly counts of lobster at each site each year, from 2012 to 2018. 

### Results  
#### A. Annual abundance of lobsters at each site
The changes in annual abundance of California Spiny Lobster at five sites were plotted to visualize (Figure 1). Comparing 2018 to 2012, spiny lobsters from all sites experienced an increasing trend in abundance except those from Arroyo Quemado, whose stock mainly stayed the same. Populations from Marine Protected Areas (MPA) both peaked in 2015 then quickly recovered from 2016 to 2018, especially for Isla Vista lobsters. However, those from non-MPAs all declined after the peak and failed to recover to the peak level in the subsequent years.

```{r}
mpa_site <- c("NAPL", "IVEE")

abdn_year_site <- lobster_abundance %>% 
  count(year, site) %>% 
  mutate(mpa = ifelse(site %in% mpa_site, "(MPA)", "(non-MPA)")) %>% 
  mutate(year_abb = year - 2000) %>% 
  mutate(site_labels = factor(site, levels = c("AQUE", "CARP", "MOHK", "IVEE", "NAPL"),
                           labels = c("Arroyo Quemado", "Carpinteria",  "Mohawk", "Isla Vista", "Naples")))

ggplot(data = abdn_year_site, 
       aes(x = year_abb,
           y = n)) +
  geom_line(aes(color = mpa), 
            show.legend = FALSE) +
  geom_point(aes(color = mpa),
             alpha = 0.5,
             show.legend = FALSE) +
  facet_wrap(site_labels ~ mpa) +
  labs(x = "Year", 
       y = "Annual abundance") +
  theme_light()

```

_**Figure 1.** The changes in annual abundance of spiny lobsters at each site (2012-2018). Sites were from both MPAs (red) and non-MPAs (blue). Data: SBC LTER_ 

--------------------

#### B. The shift of lobster size distribution at each site
Lobster size distribution presented different levels of shifts at each site comparing 2018 to 2012 (Figure 2). Overall, based on the observations, the size structure of lobster population shifted to the right from 2012 to 2018 with 2018 also had much greater population. Specifically, the shift to larger sizes was more significant in sites located in MPA. Lobsters in Isla Vista and Naples were clustered to smaller sizes in 2012 but grew larger in 2018 under protections from fishing pressure. However, for those in non-MPAs, lobsters did not grow larger if not dropped to smaller sizes.

```{r}
size_1218_site <- lobster_abundance %>% 
  filter(year %in% c("2012", "2018")) %>% 
  select(year, site, size_mm) %>% 
  mutate(mpa = ifelse(site %in% mpa_site, "(MPA)", "(non-MPA)")) %>% 
  mutate(site_labels = factor(site, levels = c("AQUE", "CARP", "MOHK", "IVEE", "NAPL"),
                           labels = c("Arroyo Quemado", "Carpinteria",  "Mohawk", "Isla Vista", "Naples")))

size_1218 <- size_1218_site %>% 
  transform(site_labels = NULL, mpa = NULL)

ggplot(data = size_1218_site, 
       aes(x = size_mm)) +
  geom_histogram(data = size_1218, 
                 fill = "gray90", 
                 binwidth = 5) +
  geom_histogram(aes(fill = mpa), 
                 binwidth = 5,
            show.legend = FALSE) +
  facet_grid(year ~ site_labels + mpa) +
  labs(x = "Size(mm)", 
       y = "Count") +
  theme_light()
```

_**Figure 2.** The shift of loster size distribution in 2012 and 2018 by site. In 2012, sites in non-MPA (blue) had more larger sizes lobsters than sites in MPA (red). In 2018, lobsters in MPAs had larger sizes than those from non-MPAs. Data: SBC LTER_ 

-----------------------

#### C. Mean lobster sizes at MPA and non-MPA sites in 2012 and 2018
```{r}
size_1218_mpa <- lobster_abundance %>% 
  filter(year %in% c("2012", "2018")) %>% 
  select(year, site, size_mm) %>% 
  mutate(mpa = ifelse(site %in% mpa_site, "MPA", "non-MPA"))

size_table <- size_1218_mpa %>% 
  group_by(year, mpa) %>% 
  summarize(mean = round(mean(size_mm), 2),
            median = round(median(size_mm), 2),
            standard_deviation = round(sd(size_mm), 2),
            sample_size = length(size_mm))
```
**a summary paragraph here**

_**MPA vs. non-MPA sites in 2012**_
```{r}
size_12 <- size_1218_mpa %>% 
  filter(year == "2012")
ttest_2012 <- t.test(size_mm ~ mpa, data = size_12)

size_12_mpa <- size_12 %>% 
  filter(mpa == "MPA")
size_12_non_mpa <- size_12 %>% 
  filter(mpa == "non-MPA")
d_12 <- cohen.d(size_12_mpa$size_mm, size_12_non_mpa$size_mm)
```
For 2012 observations, mean lobster size (mm) measured in samples collected from MPA sites (`r round(size_table$mean[1], 2)` $\pm$ `r round(size_table$standard_deviation[1], 2)`, n = `r size_table$sample_size[1]`) differed significantly from size of lobsters in non-MPA sites (`r round(size_table$mean[2], 2)` $\pm$ `r round(size_table$standard_deviation[2], 2)`, n = `r size_table$sample_size[2]`) by a two-sample t-test (t(`r floor(ttest_2012$parameter)`) = `r round(ttest_2012$statistic, 2)`, *p* = `r round(ttest_2012$p.value, 4)`, 95% CI = (`r round(ttest_2012$conf.int, 2)`)). The sample sizes for the two groups were not balanced where that for MPA group was small. With the *p*-value, there is a `r round(ttest_2012$p.value, 4)*100`% chance that we could have found two samples at least as different as ours if they were drawn from two populations with the same means. The 95% confidence interval did not include 0. The difference in mean lobster size between MPA and non-MPA sites was `r round(size_table$mean[2] - size_table$mean[1], 2)`mm, with a medium effect size (Cohen's d = `r round(d_12$estimate, 2)`). If we could have a larger sample size for MPA 2012, we might be able to further improve the analysis outcome.


_**MPA vs. non-MPA sites in 2018**_
```{r}
size_18 <- size_1218_mpa %>% 
  filter(year == "2018")
ttest_2018 <- t.test(size_mm ~ mpa, data = size_18)

size_18_mpa <- size_18 %>% 
  filter(mpa == "MPA")
size_18_non_mpa <- size_18 %>% 
  filter(mpa == "non-MPA")
d_18 <- cohen.d(size_18_mpa$size_mm, size_18_non_mpa$size_mm)
```
For 2018 observations, mean lobster size (mm) measured in samples collected from MPA sites (`r round(size_table$mean[3], 2)` $\pm$ `r round(size_table$standard_deviation[3], 2)`, n = `r size_table$sample_size[3]`) differed significantly from size of lobsters in non-MPA sites (`r round(size_table$mean[4], 2)` $\pm$ `r round(size_table$standard_deviation[4], 2)`, n = `r size_table$sample_size[4]`) by a two-sample t-test (t(`r floor(ttest_2018$parameter)`) = `r round(ttest_2018$statistic, 2)`, *p* < 0.001, 95% CI = (`r round(ttest_2018$conf.int, 2)`)). Based simply on *p*-value, there was an extremely rare chance that we could have found two samples at least as different as ours if they were drawn from two populations with the same means. Also, the 95% confidence interval did not include 0. However, the difference in mean lobster size between MPA and non-MPA sites was `r round(size_table$mean[3] - size_table$mean[4], 2)`mm, with a small effect size (Cohen's d = `r round(d_18$estimate, 2)`). The significant t-test result might be heavily influenced by large sample sizes of the two groups.


_**2012 vs. 2018 for MPA sites**_
```{r}

size_mpa <- size_1218_mpa %>% 
  filter(mpa == "MPA")
ttest_mpa <- t.test(size_mm ~ year, data = size_mpa)
ttest_mpa

size_mpa_12 <- size_mpa %>% 
  filter(year == "2012")
size_mpa_18 <- size_mpa %>% 
  filter(year == "2018")
d_mpa <- cohen.d(size_mpa_12$size_mm, size_mpa_18$size_mm)
```
For observations in MPA sites, mean lobster size (mm) measured in samples collected during 2012 (`r round(size_table$mean[1], 2)` $\pm$ `r round(size_table$standard_deviation[1], 2)`, n = `r size_table$sample_size[1]`) differed significantly from size of lobsters measured during 2018 (`r round(size_table$mean[3], 2)` $\pm$ `r round(size_table$standard_deviation[3], 2)`, n = `r size_table$sample_size[3]`) by a two-sample t-test (t(`r floor(ttest_mpa$parameter)`) = `r round(ttest_mpa$statistic, 2)`, *p* < 0.001, 95% CI = (`r round(ttest_mpa$conf.int, 2)`)). The sample sizes for the two groups were even more unbalanced than those in the first hypothesis test. With the *p*-value, there is a very small chance that we could have found two samples at least as different as ours if they were drawn from two populations with the same means. The 95% confidence interval did not include 0. The difference in mean lobster size between MPA and non-MPA sites was `r round(size_table$mean[3] - size_table$mean[1], 2)`mm, with a large effect size (Cohen's d = `r round(d_mpa$estimate, 2)`). The large effect size supported the significant difference between lobster sizes measured in 2012 and 2018 in MPA sites, even when one sample size was much smaller than the other. The result gave strong evidence that spiny lobster in MPA sites recovered effectively under protections against fishing pressure.

```{r}
size_non_mpa <- size_1218_mpa %>% 
  filter(mpa == "non-MPA")
ttest_non_mpa <- t.test(size_mm ~ year, data = size_non_mpa)
ttest_non_mpa

size_non_mpa_12 <- size_non_mpa %>% 
  filter(year == "2012")
size_non_mpa_18 <- size_non_mpa %>% 
  filter(year == "2018")
d_non_mpa <- cohen.d(size_non_mpa_12$size_mm, size_non_mpa_18$size_mm)
```

_**Table 1.** Summary of spiny lobster size for MPA and non-MPA sites in 2012 and 2018. For non-MPA sites, lobster size did not change much from 2012 (mean = `r size_table$mean[2]` , sd = `r size_table$standard_deviation[2]`, n = `r size_table$sample_size[2]`) to 2018 (mean = `r size_table$mean[4]` , sd = `r size_table$standard_deviation[4]`, n = `r size_table$sample_size[4]`). For MPA sites, lobster size increased between 2012 (mean = `r size_table$mean[1]` , sd = `r size_table$standard_deviation[1]`, n = `r size_table$sample_size[1]`) and 2018 (mean = `r size_table$mean[3]` , sd = `r size_table$standard_deviation[3]`, n = `r size_table$sample_size[3]`). Data: SBC LTER_ 
```{r}
size_table %>% 
  kable(col.names = c("Year", "MPA", "Mean", "Median", "Standard Deviation", "Sample Size")) %>% 
  kable_styling(bootstrap_options = "striped",
                position = "center",
                full_width = FALSE)
```

---------------------

### Summary

### References

